\\fontsize{12pt}{3pt}\\selectfont \\textbf{\\textit{Submitted to Ecosphere. Manuscript type: Article}}

### 2023-10-19 manuscript/manuscriptRW.Rmd
change figures names and remove them from manuscript

Recruitment data covering a wide range of environmental conditions are
hard to obtain, and this is one of the reasons why most models of forest
dynamics have never been confronted with a dataset covering such
gradients over a large number of sites, to evaluate how well
regeneration is captured. The observations used here are derived from a
novel network of sites in forest reserves that represent the range of
environmental gradients in temperature and precipitation in Central
Europe as compiled in the framework of the EuFoRIa network
[@euforia2019; @käber2023] (Figure \@ref(fig:map)). The environmental
coverage and the variety of forests represented by this data is
unprecedented within Europe. These forests have been unmanaged for at
least 10 years prior to their designation as reserves, and most of them
provide time series of natural forest dynamics over multiple decades and
up to ca. 80 years. The census periods range from 3 to 37 years, with an
average of 14 years. The data provide information at the tree level,
thus allowing for the sequential comparison of processes such as
individual-tree regeneration and death between the measurements.

@@

Simulated levels of recruitment varied strongly across the 15 models and
typically did not match the levels found in the forest reserve data
(Figure \@ref(fig:over)). Recruitment was overestimated in most models
for both the 7 and 10 cm diameter thresholds, with the exception of the
empirical stand model SIBYLA, the landscape model Landis-II and the
global model aDGVM2, which estimated recruited levels at the lower end
of the plausibility interval of the observed data (Figure
\@ref(fig:over)). The models with the largest overestimation were the
stand model PICUS and the landscape model TreeMig. For most models, the
variability of simulated recruitment levels across the 200 sites
(visible from the interquartile range in the box plots of Figure
\@ref(fig:over)) was similar to or smaller than observed, with the
notable exception of PICUS, where simulated recruitment variability was
much larger.

@@

Most models matched the level of diversity of the observed data quite
well (Figure \@ref(fig:H710)). Five models overestimated recruitment
diversity: ForCEEPS, ForClim 1, PICUS, TreeMig and LandClim, the latter
particularly for the 7 cm diameter threshold. The model 4C is a special
case, as it simulated five species only, i.e., its diversity values are
not directly comparable to those of the other models, nor to the
observations. Only one model, Landis-II, consistently underestimated
recruited diversity.

@@

For both the observed and the simulated data, and both recruitment
thresholds, species diversity in the recruitment compared to stand-level
diversity did not feature clear patterns across the models (Figures S1
and S2). Most models captured reasonably well or overpredicted species
diversity at the stand level (Figure \@ref(fig:combinedRichness)A, left
and center; cf. Figure S3), but a characteristic feature was that the
simulations had a much lower variability of diversity compared to
observations. Models overpredicting species diversity included ForClim1,
ForClim 11, FORMIND and TreeMig (Figure \@ref(fig:combinedRichness)A,
left). Several models underpredicted species diversity, i.e., LPJ-GUESS,
ForCEEPS(f) and 4C (Figure \@ref(fig:combinedRichness)A, right).

The majority of the models overpredicted recruitment diversity; as
already observed in Figure \@ref(fig:H710), only two of them
underestimated it (4C and Landis-II) (Figure
\@ref(fig:combinedRichness)B and Figure S4). Both the models with
feedback and those without feedback from the adult trees to regeneration
(via seed production) overestimated, underestimated or captured
reasonably well recruitment species diversity compared to the
observations. Again, simulated diversity had considerably lower
variability than observed diversity in the recruitment.

@@

There were strong differences among the models regarding the mortality
rate between 7 and 10 cm (Figure \@ref(fig:mortality11)). Most of the
models featured a mortality rate significantly larger than the observed
data, thus at least partially compensating for the general
overestimation of recruitment levels (Figure \@ref(fig:over) and Table
S3); six models did not feature a significant difference compared to the
observed data (ForCEEPS, FORMIND, ForClim 1 and 11, iLand and aDGVM2).
Four models (xComp, PICUS, LandClim and TreeMig) featured very high
mortality rates (i.e., well above the Reineke self-thinning line), which
compensated for the strong initial overestimation of recruitment (cf.
Figure \@ref(fig:over)). Interestingly, two models that underestimated
overall recruitment levels (SIBYLA and Landis-II; cf. Figure
\@ref(fig:over)) featured mortality rates that were close to but still
above the self-thinning line (Figure \@ref(fig:mortality11)).

@@

The three models with the largest overestimation of the proportion of
recruitment at the 7 cm threshold featured the highest mortality between
the 7 and 10 cm thresholds (Figure \@ref(fig:mortality13)A). The other
models that overestimated recruitment had mortalities lower than the
expected self-thinning ratio, with the exception of LandClim (Figure
\@ref(fig:mortality13)B). It is noteworthy that most of the models that
featured a low ratio of regeneration between 7 and 10 cm (i.e.,
ForCEEPS, ForClim 1, ForClim 11 and iLand) had only a small
overestimation of recruitment at the 7 cm threshold (Figure
\@ref(fig:mortality11)). aDGVM2 was the only model that underestimated
recruitment at the 7 cm threshold while featuring a low mortality
between the two thresholds (Figure \@ref(fig:mortality13)C), and only a
few models were close to the correct recruitment levels at 7 cm while
also being close to the Reineke line (SIBYLA, Landis-II; cf. Figure
\@ref(fig:mortality13)D; and 4C as well as LandClim).

@@

When evaluating total recruited levels along key gradients of light
availability (basal area), temperature (degree-days) and soil moisture
(climatic water balance), distinct features emerged: the models
reproduced the effect of basal area in both its magnitude and patterns
(Figure \@ref(fig:nvTrend710BAlightTemp) and S7) considerably better
than the effects of the climatic gradients, where they featured varying
patterns.

In the observed data, recruited levels decreased clearly with increasing
stand basal area (Figure \@ref(fig:nvTrend710BAlightTemp)a). This trend
was captured by the Landis-II model only, albeit at much lower values of
basal area than in the observations. The other models featured
distinctly different trends, such as (1) an increase in recruited levels
with increasing total basal area followed by the absence of regeneration
at high values of basal area (ForCEEPS, ForCEEPS(f), FORMIND, TreeMig
and LPJ-GUESS), (2) almost constant recruitment levels with basal area
(ForClim 11, xComp, LandClim and aDGVM2), or (3) an increase of
recruitment up to a certain value of stand basal area followed by a
decrease at even higher values of stand basal area, with model-specific
thresholds (4C, ForClim 1, PICUS, SIBYLA and iLand).

The observed recruitment did not change much across the climatic water
balance gradient, and showed a slight increase with temperature. The
models did not match the observed recruitment patterns across the
gradients of soil moisture and temperature (Figure
\@ref(fig:nvTrend710BAlightTemp)b,c), and neither was there a clear
pattern across models. Regarding the soil moisture gradient, a group of
models featured decreasing recruitment with increasing soil moisture
(4C, ForClim 1, ForClim11, SIBYLA, xComp, PICUS, LandClim, TreeMig and
iLand), while a few models showed the opposite trend (ForCEEPS(f),
FORMIND and LPJ-GUESS). Across the temperature gradient, all the models
that featured a decrease in recruitment levels with increasing soil
moisture showed the opposite trend (i.e., more recruitment with
increasing temperature), except for 4C. The other models did not feature
a clear pattern.

Silver fir (*Abies alba*), beech (*Fagus sylvatica*), Scots pine (*Pinus
sylvestris*) and oak (*Quercus* spp.) were the species observed most
frequently in the EuFoRIa data. Most models captured well the share of
basal area in the recruitment of these five species compared to the
observed data along the two major climatic gradients (Figure
\@ref(fig:diff), cf. Figure S8). However, the models differed
considerably in the way in which they simulated these climatic niches.

First, some models overestimated recruitment across the entire gradients
of at least one of the five main tree species (Figure \@ref(fig:diff),
e.g., ForCEEPS, ForCEEPS(f), LandClim or LPJ-GUESS), while other models
overestimated recruitment of more than one species for a large part of
the climate gradients, such as xComp, PICUS, SIBYLA or LPJ-GUESS (cf.
red colors in Figure \@ref(fig:diff)). These trends were consistent for
both recruitment thresholds, although the exact changes in the share of
the recruited basal area were different (Figure S8).

@@

Lastly, there is no model that performed well across the five species
regarding the species-specific recruitment levels in the environmental
space (Figure \@ref(fig:diff)). Some models (e.g., ForClim11, PICUS)
tended to consistently overestimate the recruitment share of all five
species, but most models overestimated the recruitment share of some
species while underestimating the share of others (e.g., ForCEEPS,
LPJ-GUESS).

@@

```{r map, echo=FALSE, fig.cap=".", out.width='80%'}
knitr::include_graphics(here::here("figures", "theMap.png"))
```

```{r over, echo=FALSE, fig.cap=".", out.width='100%'}
        knitr::include_graphics(here::here("figures", "recOverAll710.png"))
```

```{r H710, echo=FALSE, fig.cap=".", out.width='100%'}
    knitr::include_graphics(here::here("figures", "H7_10.png"))
```

```{r combinedRichness, echo=FALSE, fig.cap=".", out.width='75%'}
knitr::include_graphics(here::here("figures", "combined_richness.png"))
```

```{r mortality11, echo=FALSE, fig.cap=".", out.width='75%'}
knitr::include_graphics(here::here("figures", "ratio7_10.png"))
```

```{r mortality13, echo=FALSE, fig.cap=".", out.width='80%'}
knitr::include_graphics(here::here("figures", "mort7_10_medianOver.png"))
```

```{r nvTrend710BAlightTemp, echo=FALSE, fig.cap=".", fig.show = 'hold', out.width="49%",fig.ncol = 2, fig.subcap=c('', '', ''), fig.show='hold', fig.align='left'}

   knitr::include_graphics(c(here::here("figures", "envTrend7Totba.png"),
                              here::here("figures", "envTrend7WB.png"),
                              here::here("figures", "envTrend7DDS.png")))
```

>>>>

Recruitment data covering a wide range of environmental conditions are
hard to obtain, and this is one of the reasons why most models of forest
dynamics have never been confronted with a dataset covering such
gradients over a large number of sites, to evaluate how well
regeneration is captured. The observations used here are derived from a
novel network of sites in forest reserves that represent the range of
environmental gradients in temperature and precipitation in Central
Europe as compiled in the framework of the EuFoRIa network
[@euforia2019; @käber2023] (Figure 1). The environmental coverage and
the variety of forests represented by this data is unprecedented within
Europe. These forests have been unmanaged for at least 10 years prior to
their designation as reserves, and most of them provide time series of
natural forest dynamics over multiple decades and up to ca. 80 years.
The census periods range from 3 to 37 years, with an average of 14
years. The data provide information at the tree level, thus allowing for
the sequential comparison of processes such as individual-tree
regeneration and death between the measurements.

@@

Simulated levels of recruitment varied strongly across the 15 models and
typically did not match the levels found in the forest reserve data
(Figure 2). Recruitment was overestimated in most models for both the 7
and 10 cm diameter thresholds, with the exception of the empirical stand
model SIBYLA, the landscape model Landis-II and the global model aDGVM2,
which estimated recruited levels at the lower end of the plausibility
interval of the observed data (Figure 2). The models with the largest
overestimation were the stand model PICUS and the landscape model
TreeMig. For most models, the variability of simulated recruitment
levels across the 200 sites (visible from the interquartile range in the
box plots of Figure 2) was similar to or smaller than observed, with the
notable exception of PICUS, where simulated recruitment variability was
much larger.

@@

Most models matched the level of diversity of the observed data quite
well (Figure 3). Five models overestimated recruitment diversity:
ForCEEPS, ForClim 1, PICUS, TreeMig and LandClim, the latter
particularly for the 7 cm diameter threshold. The model 4C is a special
case, as it simulated five species only, i.e., its diversity values are
not directly comparable to those of the other models, nor to the
observations. Only one model, Landis-II, consistently underestimated
recruited diversity.

@@

For both the observed and the simulated data, and both recruitment
thresholds, species diversity in the recruitment compared to stand-level
diversity did not feature clear patterns across the models (Figures S1
and S2). Most models captured reasonably well or overpredicted species
diversity at the stand level (Figure 4A, left and center; cf. Figure
S3), but a characteristic feature was that the simulations had a much
lower variability of diversity compared to observations. Models
overpredicting species diversity included ForClim1, ForClim 11, FORMIND
and TreeMig (Figure 4A, left). Several models underpredicted species
diversity, i.e., LPJ-GUESS, ForCEEPS(f) and 4C (Figure 4A, right).

The majority of the models overpredicted recruitment diversity; as
already observed in Figure 3, only two of them underestimated it (4C and
Landis-II) (Figure 4B and Figure S4). Both the models with feedback and
those without feedback from the adult trees to regeneration (via seed
production) overestimated, underestimated or captured reasonably well
recruitment species diversity compared to the observations. Again,
simulated diversity had considerably lower variability than observed
diversity in the recruitment.

@@

There were strong differences among the models regarding the mortality
rate between 7 and 10 cm (Figure 5). Most of the models featured a
mortality rate significantly larger than the observed data, thus at
least partially compensating for the general overestimation of
recruitment levels (Figure 2 and Table S3); six models did not feature a
significant difference compared to the observed data (ForCEEPS, FORMIND,
ForClim 1 and 11, iLand and aDGVM2). Four models (xComp, PICUS, LandClim
and TreeMig) featured very high mortality rates (i.e., well above the
Reineke self-thinning line), which compensated for the strong initial
overestimation of recruitment (cf. Figure 2). Interestingly, two models
that underestimated overall recruitment levels (SIBYLA and Landis-II;
cf. Figure 2) featured mortality rates that were close to but still
above the self-thinning line (Figure 5).

@@

The three models with the largest overestimation of the proportion of
recruitment at the 7 cm threshold featured the highest mortality between
the 7 and 10 cm thresholds (Figure 6A). The other models that
overestimated recruitment had mortalities lower than the expected
self-thinning ratio, with the exception of LandClim (Figure 6B). It is
noteworthy that most of the models that featured a low ratio of
regeneration between 7 and 10 cm (i.e., ForCEEPS, ForClim 1, ForClim 11
and iLand) had only a small overestimation of recruitment at the 7 cm
threshold (Figure 5). aDGVM2 was the only model that underestimated
recruitment at the 7 cm threshold while featuring a low mortality
between the two thresholds (Figure 6C), and only a few models were close
to the correct recruitment levels at 7 cm while also being close to the
Reineke line (SIBYLA, Landis-II; cf. Figure 6D; and 4C as well as
LandClim).

@@

When evaluating total recruited levels along key gradients of light
availability (basal area), temperature (degree-days) and soil moisture
(climatic water balance), distinct features emerged: the models
reproduced the effect of basal area in both its magnitude and patterns
(Figure 7 and S7) considerably better than the effects of the climatic
gradients, where they featured varying patterns.

In the observed data, recruited levels decreased clearly with increasing
stand basal area (Figure 7a). This trend was captured by the Landis-II
model only, albeit at much lower values of basal area than in the
observations. The other models featured distinctly different trends,
such as (1) an increase in recruited levels with increasing total basal
area followed by the absence of regeneration at high values of basal
area (ForCEEPS, ForCEEPS(f), FORMIND, TreeMig and LPJ-GUESS), (2) almost
constant recruitment levels with basal area (ForClim 11, xComp, LandClim
and aDGVM2), or (3) an increase of recruitment up to a certain value of
stand basal area followed by a decrease at even higher values of stand
basal area, with model-specific thresholds (4C, ForClim 1, PICUS, SIBYLA
and iLand).

The observed recruitment did not change much across the climatic water
balance gradient, and showed a slight increase with temperature. The
models did not match the observed recruitment patterns across the
gradients of soil moisture and temperature (Figure 7b,c), and neither
was there a clear pattern across models. Regarding the soil moisture
gradient, a group of models featured decreasing recruitment with
increasing soil moisture (4C, ForClim 1, ForClim11, SIBYLA, xComp,
PICUS, LandClim, TreeMig and iLand), while a few models showed the
opposite trend (ForCEEPS(f), FORMIND and LPJ-GUESS). Across the
temperature gradient, all the models that featured a decrease in
recruitment levels with increasing soil moisture showed the opposite
trend (i.e., more recruitment with increasing temperature), except for
4C. The other models did not feature a clear pattern.

Silver fir (*Abies alba*), beech (*Fagus sylvatica*), Scots pine (*Pinus
sylvestris*) and oak (*Quercus* spp.) were the species observed most
frequently in the EuFoRIa data. Most models captured well the share of
basal area in the recruitment of these five species compared to the
observed data along the two major climatic gradients (Figure 8, cf.
Figure S8). However, the models differed considerably in the way in
which they simulated these climatic niches.

First, some models overestimated recruitment across the entire gradients
of at least one of the five main tree species (Figure 8, e.g., ForCEEPS,
ForCEEPS(f), LandClim or LPJ-GUESS), while other models overestimated
recruitment of more than one species for a large part of the climate
gradients, such as xComp, PICUS, SIBYLA or LPJ-GUESS (cf. red colors in
Figure 8). These trends were consistent for both recruitment thresholds,
although the exact changes in the share of the recruited basal area were
different (Figure S8).

@@

Lastly, there is no model that performed well across the five species
regarding the species-specific recruitment levels in the environmental
space (Figure 8). Some models (e.g., ForClim11, PICUS) tended to
consistently overestimate the recruitment share of all five species, but
most models overestimated the recruitment share of some species while
underestimating the share of others (e.g., ForCEEPS, LPJ-GUESS).

@@
